Figure two

### 2025-10-27 draftFiguretwo_Ella.Rmd
minor changes Fig 2 (color scale surveys)

# Define which packages needed for analyses
p_needed <-
  c("knitr",
    "dplyr", 
    "stargazer",
    "ggplot2",
    "viridis",
    "haven",
    "MASS",
    "modelsummary",
    "tidyr",
    "data.table",
    "scales")

@@

## Clean the data (would skip this?)

@@

#-------------------------------------------------------------------------------
# consistent country names 
#-------------------------------------------------------------------------------

## paper
papers <- papers %>%
  mutate(country_clean = recode(country,
    # Turkey variants
    "Turkey" = "Turkyie",
    "Türkiye" = "Turkyie",

    # UK variants
    "UK" = "United Kingdom",
    "England" = "United Kingdom",
    "Wales" = "United Kingdom",

    # Korea variants
    "Korea" = "South Korea",

    # Netherlands variants
    "The Netherlands" = "Netherlands",

    # Czechia variants
    "Czech Republic" = "Czechia",

    # US variants
    "USA" = "United States",

    .default = country  # keep as is otherwise
  ))

papers <- papers[which(papers$country_clean != "exclude"),]
papers <- papers[which(papers$country_clean != "no info"),]

## survey
surveys <- surveys %>%
  mutate(country_clean = recode(countries,
    # Turkey variants
    "Turkey" = "Turkyie",
    "Türkiye" = "Turkyie",

    # UK variants
    "UK" = "United Kingdom",
    "England" = "United Kingdom",
    "Wales" = "United Kingdom",

    # Korea variants
    "Korea" = "South Korea",

    # Netherlands variants
    "The Netherlands" = "Netherlands",

    # Czechia variants
    "Czech Republic" = "Czechia",

    # US variants
    "USA" = "United States",

    .default = countries  # keep as is otherwise
  ))

#-------------------------------------------------------------------------------
# relevant columns
#-------------------------------------------------------------------------------
papers <- papers[,c("ID", "country_clean",
                    "Income.group", "Region.x", 
                    "Perceptions", "Behavior", "Priority", "Policy", "Health", "None", "pm25Avg")]

surveys <- surveys[,c("id", "country_clean", "year", "Income.group", "Region.x", 
                      "Perceptions", "Behavior", "Priority", "Policy", "Health", "Inclusion", "pm25Avg")]

@@

# install.packages("treemapify")
library(tidyverse)
library(scales)
library(treemapify)

@@

surveys_long <- surveys %>%
  count(category, name = "n") %>%   # counts per category
  mutate(share = n / sum(n)) %>%    # sum(n) is the grand total here
  rename(Topic = category)

@@

# install.packages("ggnewscale")
library(tidyverse)
library(scales)
library(ggnewscale)

@@

#"#F3EBF8", "#D4ABEF", "#541F78"

@@

#-------------------------------------------------------------------------------
# get exposure groups for papers and surveys
#-------------------------------------------------------------------------------

## now is already in the papers dataset, could change values based on pollution value (pm25$categoryPollution <- cut(papers$pm25Avg, breaks = c(0, 10, 15, 25, 35, 100), 
#                                 labels = c("Very low", "Low", "Medium", "High", "Very high"), 
#                                 right = FALSE)

# surveys$categoryPollution <- cut(surveys$pm25Avg, breaks = c(0, 10, 15, 25, 35, 100), 
#                                 labels = c("Very low", "Low", "Medium", "High", "Very high"), 
#                                 right = FALSE)

@@

  # --- main topic rows ---
  geom_tile(data = topics_data_poll_s,
            aes(x = categoryPollution, y = Topic, fill = share),
            color = "white", linewidth = 0.8) +
  geom_text(
    data = topics_data_poll_s,
    aes(x = categoryPollution, y = Topic,
        label = scales::percent(share, accuracy = 1),
        color = ifelse(share > 0.25, "white", "black")),
    size = 4) +
  scale_fill_gradient(
    name = "Share within\nexposure group",
    low = "#f7f4f9", high = "#8B475D",
    labels = scales::percent,
    breaks = c(min(topics_data_poll_s$share), max(topics_data_poll_s$share))) +
  scale_color_identity() +
labs(x = "Exposure group", y = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1, size = 13),
    axis.text.y = element_text(size = 13),
    legend.position = "top")

>>>>

# Define which packages needed for analyses
p_needed <-
  c("knitr",
    "dplyr", 
    "stargazer",
    "ggplot2",
    "viridis",
    "haven",
    "MASS",
    "modelsummary",
    "tidyr",
    "data.table",
    "scales",
    "tidyverse",
    "treemapify",
    "ggnewscale")

@@

## Filter (only Inclusion = 1 items)

@@

#-------------------------------------------------------------------------------
# Inclusion = 1 for survey items 
#-------------------------------------------------------------------------------

@@

@@

#-------------------------------------------------------------------------------
# surveys
#-------------------------------------------------------------------------------
surveys_long <- surveys %>%
  count(category, name = "n") %>%   # counts per category
  mutate(share = n / sum(n)) %>%    # sum(n) is the grand total here
  rename(Topic = category)

@@

@@

@@

@@

  # --- main topic rows ---
  geom_tile(data = topics_data_poll_s,
            aes(x = categoryPollution, y = Topic, fill = share),
            color = "white", linewidth = 0.8) +
  geom_text(
    data = topics_data_poll_s,
    aes(x = categoryPollution, y = Topic,
        label = scales::percent(share, accuracy = 1),
        color = ifelse(share > 0.25, "white", "black")),
    size = 4) +
  scale_fill_gradient(
    name = "Share within\nexposure group",
    low = "#F3EBF8", high = "#541F78",
    labels = scales::percent,
    breaks = c(min(topics_data_poll_s$share), max(topics_data_poll_s$share))) +
  scale_color_identity() +
labs(x = "Exposure group", y = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1, size = 13),
    axis.text.y = element_text(size = 13),
    legend.position = "top")
